The population of long-period exoplanets --- \\ Fully-automated discovery \& characterization of long-period transiting exoplanet candidates in the \kepler\ archive \vspace{-3\baselineskip}

### 2016-05-14 document/ms.tex
writing

% This paper is part of the single transits project.
% Copyright 2015 Dan Foreman-Mackey (NYU) and the co-authors listed below.
%
%  RULES OF THE GAME
%
%  * 80 characters
%  * line breaks at the ends of sentences
%  * eqnarrys ONLY
%  * ``light curve'' not ``light-curve'' or ``lightcurve''
%  * that is all.
%

% \documentclass[twocolumn]{aastex6}
% \documentclass[onecolumn]{aastex6}
\documentclass[preprint,12pt]{aastex}

@@

\usepackage{color,hyperref}
\definecolor{linkcolor}{rgb}{0,0,0.5}
\hypersetup{colorlinks=true,linkcolor=linkcolor,citecolor=linkcolor,
            filecolor=linkcolor,urlcolor=linkcolor}
\usepackage{url}
\usepackage{amssymb,amsmath}
\usepackage{subfigure}
\usepackage{booktabs}

\usepackage{natbib}
\bibliographystyle{aasjournal}

\newcommand{\project}[1]{\textsl{#1}}
\newcommand{\kepler}{\project{Kepler}}
\newcommand{\KT}{\project{K2}}
\newcommand{\tess}{\project{TESS}}
\newcommand{\plato}{\project{PLATO}}
\newcommand{\jwst}{\project{JWST}}
\newcommand{\terra}{\project{TERRA}}
\newcommand{\pdc}{\project{PDC}}
\newcommand{\license}{MIT License}

\newcommand{\paper}{article}

@@

\newcommand{\datareleaseurl}{{\url{http://bbq.dfm.io/ketu}}}

@@

% \slugcomment{Not to appear in Nonlearned J., 45.}

@@

\title{%
The population of long-period exoplanets --- \\
Fully-automated discovery \& characterization of
long-period transiting exoplanet candidates in the \kepler\ archive
}

@@

The transit method of exoplanet detection and characterization has been
demonstrated as a powerful method for building systematic catalogs of
exoplanets.
Despite the great success of the \kepler\ Mission with thousands of planet
discoveries \citep{Burke:2014, Rowe:2015}, current methods for exoplanet
discoveries are currently limited in the range of orbital periods that can be
studied.
Specifically, the standard transit search procedures only discover signals
with at least three observed transits \citep[for example][]{Petigura:2013,
Burke:2014, Rowe:2015}.
For \kepler, with a baseline of about four years, this sets an absolute upper
limit on the detectable periods of about two years.
In the Solar System, Jupiter~--~with a period of 12 years~--~dominates the
planetary dynamics and, since it would only exhibit at most one transit in the
\kepler\ data, it would be missed by most existing transit search procedures.
It is possible to discover long-period planets like this using targeted radial
velocity (RV) surveys \citep[for example][]{Butler:2006, Knutson:2014}
\dfmtodo{CITE Bryan \etal} but the cost of implementing a systematic RV search
is substantially higher than searching the existing and forthcoming
photometric data for single transits.

There are two main technical barriers to a search for single transit events.
The first is that the transit probability for long-period planets is very low;
scaling as $\propto P^{-5/3}$ for orbital periods longer than the
baseline of contiguous observations.
Therefore, even if long-period planets are intrinsically common, they will
still be underrepresented in a transiting sample.
The second challenge is that there are many signals in the observed
light curves caused by stochastic processes~--~both instrumental
(pointing jitter, temperature variations, \etc) and astrophysical (stellar
variability, \etc)~--~that can masquerade as transits.
Even using the most sophisticated methods for removing this variability, false
signal far outnumber the true single transits.

Nearly every transit search algorithm is built on the same principles and many
of the same decisions are made.
In particular, at the heart of most methods is a matched filtering step where
the likelihood of an approximate transit model is computed on a grid in the
physical parameters \citep[\kepler\ Data Processing
Handbook\footnote{\url{https://archive.stsci.edu/kepler/manuals/KSCI-19081-001_Data_Processing_Handbook.pdf}};][]{%
Petigura:2013, Huang:2013, Dressing:2015, Foreman-Mackey:2015}.
Using these methods, and substantial hand-curation, some long-period
transiting candidates have been published \citep[for example][]{Batalha:2013,
Huang:2013, Kipping:2014a}.
For more recent data releases, the community has settled on more conservative
selection criteria where candidates are required to have multiple transits
\citep[for example][]{Petigura:2013, Burke:2014, Rowe:2015}.
Even the \project{QATS} algorithm \citep{Carter:2013} for finding
quasiperiodic transits builds on much the same infrastructure.
This means that there has never been a systematic search for single transits
in the full \kepler\ dataset.

A qualitatively different approach to planet search is employed by the
\project{Planet Hunters} (PH)
project\footnote{\url{http://www.planethunters.org/}} \citep{Fischer:2012}.
PH is a ``citizen science'' project where visitors to the website look at
sections of light curve and mark the locations of transits that can be
identified visually.
Visual inspection can be a useful search technique for large single transits
because humans are able to robustly distinguish transit signals from the noise
and this project has yielded some promising long-period candidates and
confirmed planets \citep[for example][]{Wang:2013} \dfmtodo{CITE Wang 2015}.
One shortcoming of the PH method is that it can be difficult to fully quantify
the performance (completeness and reliability) of the search results.
% and PH does not
% evaluate their users using synthetic transit signals as is now common practice
% in the transit search literature \dfmtodo{is this still true?}.
This means that the PH sample of candidates cannot be reliably used for robust
population inference.

Visual inspection and hand-curation by experts has also been used to find the
transits of long-period planets in previously known planetary systems
(\dfmtodo{CITE single transit KOIs}).
All light curves in the \kepler\ archive with \kepler\ Objects of Interest
(KOIs) were visually inspected to find candidate single transits.
This is a reasonable idea because long-period planets with transiting inner
planets are expected to have a higher geometric transit probability.
The selection effects introduced by restricting the target sample like this
are difficult to take into account when making inferences about the population
of long-period planets.

To study the occurrence rate of long-period exoplanets, we have developed a
fully-automated method to search the archival \kepler\ light curves and
reliably discover long-period planets with only one or two transits.
Using this method, we search the 40,000 brightest Sun-like stars and detect
XXX long-period astrophysical transit candidates.
Of these, YYY are likely to be exoplanets with radii equal to or smaller than
Jupiter's radius.
Using transit injection and recovery, we measure the completeness of our
search procedure.
We estimate that the population level astrophysical false positive rate in
this candidate sample is about ZZZ\%.
By injecting synthetic transit signals into the \kepler\ light curves, we
measure the completeness of our search method.
Based on this catalog of exoplanet candidates and the search completeness, we
constrain the occurrence rate of long-period exoplanets.
For large planets, these results are consistent with previous measurements of
long-period, giant planet occurrence rates based on radial velocity surveys
and we demonstrate that small~--~and possibly rocky~--~planets are more common
than their larger counterparts even at long orbital periods.

@@

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{figures/model_comp.pdf}
\end{center}
\caption{%
Representative examples of candidate events flagged by the initial search.
Each example falls into a different model category and the figure shows the
data as black points and the best fit mean model prediction.
The examples represent the following model categories:
\emph{(a)} variability, \emph{(b)} step, \emph{(c)} box, and \emph{(d)}
transit.
\figlabel{model-comp}}
\end{figure*}

@@

\begin{figure*}
\begin{center}
\includegraphics{figures/targets.pdf}
\end{center}
\caption{%
Blah.
\figlabel{targets}}
\end{figure*}

@@

\begin{figure*}
\begin{center}
\includegraphics{figures/full_sample.pdf}
\end{center}
\caption{%
Blah.
\figlabel{full-sample}}
\end{figure*}

@@

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{figures/completeness.pdf}
\end{center}
\caption{%
\figlabel{completeness}}
\end{figure*}

@@

\acknowledgments
It is a pleasure to thank
\ldots
for helpful contributions to the ideas and code presented here.
DFM and DWH were partially supported by the National Science Foundation
(grant IIS-1124794),
the National Aeronautics and Space Administration
(grant NNX12AI50G), and the Moore--Sloan Data Science Environment at NYU.

This research made use of the NASA \project{Astrophysics Data System} and the
NASA Exoplanet Archive.
The Archive is operated by the California Institute of Technology, under
contract with NASA under the Exoplanet Exploration Program.
This \paper\ includes data collected by the \kepler\ mission. Funding for the
\kepler\ mission is provided by the NASA Science Mission directorate.
We are grateful to the entire \kepler\ team, past and present.
Their tireless efforts were all essential to the tremendous success of the mission
and the successes of \KT, present and future.
These data were obtained from the Mikulski Archive for Space Telescopes
(MAST).
STScI is operated by the Association of Universities for Research in
Astronomy, Inc., under NASA contract NAS5-26555.
Support for MAST is provided by the NASA Office of Space Science via grant
NNX13AC07G and by other grants and contracts.
Computing resources were provided by High Performance Computing (HPC) at New
York University (NYU).

\facility{Kepler}

\appendix

@@

\clearpage
\bibliography{peerless}
\clearpage

% \begin{figure}[p]
% \begin{center}
% \includegraphics{figures/pca.pdf}
% \end{center}
% \caption{%
% The top 10 eigen light curves (ELCs) generated by running principal component
% analysis on all the aperture photometry from Campaign~1.
% \figlabel{pca}}
% \end{figure}

>>>>

% Copyright 2015-2016 Dan Foreman-Mackey and the co-authors listed below.

\documentclass[manuscript, letterpaper]{aastex6}

@@

\usepackage{url}
\usepackage{amssymb,amsmath}
\usepackage{natbib}
\bibliographystyle{aasjournal}

% omg AASTEX 6.0 -- modifications from HOGG
\setlength{\voffset}{0in}
\setlength{\hoffset}{0in}
\setlength{\textwidth}{6.5in}
\setlength{\textheight}{9in}
\setlength{\headheight}{0ex}
\setlength{\headsep}{\baselinestretch\baselineskip} % this is 2 lines in ``manuscript''
\setlength{\topmargin}{-\headsep}

\setlength{\skip\footins}{24pt}
\linespread{0.54} % close to 10/13 spacing in ``manuscript''
\setlength{\parindent}{0.54\baselineskip}
\hypersetup{colorlinks = false}
\makeatletter % you know you are living your life wrong when you need to do this
\long\def\frontmatter@title@above{
\vspace*{-\headsep}\vspace*{\headheight}
\noindent\footnotesize
{\noindent\footnotesize\textsc{\@journalinfo}}\par
{\noindent\scriptsize Preprint typeset using \LaTeX\ style AASTeX6 with
modifications by David W. Hogg and Daniel Foreman-Mackey.
}\par\vspace*{-\baselineskip}\vspace*{0.625in}
}%
\makeatother

\newcommand{\project}[1]{\textsl{#1}}

\newcommand{\kepler}{\project{Kepler}}
\newcommand{\KT}{\project{K2}}
\newcommand{\tess}{\project{TESS}}
\newcommand{\plato}{\project{PLATO}}
\newcommand{\jwst}{\project{JWST}}
\newcommand{\terra}{\project{TERRA}}
\newcommand{\pdc}{\project{PDC}}
\newcommand{\bls}{\project{BLS}}

\newcommand{\license}{MIT License}

\newcommand{\paper}{paper}

@@

\shorttitle{The population of long-period exoplanets}
\shortauthors{Foreman-Mackey, Hogg, Morton, \etal}
% \submitted{Submitted to \textit{The Astrophysical Journal}}

@@

@@

\title{%
The population of long-period exoplanets --- \\
Fully-automated discovery \& characterization of
long-period transiting exoplanet candidates in the \kepler\ archive
\vspace{-3\baselineskip}  % OMG AASTEX6 IS SO BROKEN
}

@@

Data from the \kepler\ Mission has been used to discover thousands of
transiting exoplanets.
The systematic nature of these discoveries and careful quantification of
survey selection effects, search completeness, and catalog reliability has
enabled many diverse studies of the detailed frequency and distribution of
exoplanets \citep[for example,][]{Howard:2012, Petigura:2013,
Foreman-Mackey:2014, Dressing:2015, Burke:2015, Mulders:2015}.
So far, these results have been limited to relatively short orbital periods
because existing transit search methods require the observation of multiple
transits within the baseline of the data.
For \kepler, with a baseline of about four years, this sets an absolute upper
limit on the detectable periods of about two years.
In the Solar System, Jupiter~--~with a period of 12 years~--~dominates the
planetary dynamics and, since it would only exhibit at most one transit in the
\kepler\ data, an exo-Jupiter would be missed by most existing transit search
procedures.
It is possible to discover long-period planets like this using targeted radial
velocity (RV) surveys \citep[for example][]{Butler:2006, Knutson:2014,
Bryan:2016} but the cost of implementing a systematic RV search is
substantially higher than searching the existing and forthcoming photometric
data for single transits.

There are two main technical barriers to a systematic search for single
transit events.
The first is that the transit probability for long-period planets is very low;
scaling as $\propto P^{-5/3}$ for orbital periods longer than the
baseline of contiguous observations.
Therefore, even if long-period planets are intrinsically common, they will
be underrepresented in a transiting sample.
The second challenge is that there are many signals in the observed light
curves caused by stochastic processes~--~both instrumental and
astrophysical~--~that can masquerade as transits.
Even when the most sophisticated methods for removing this variability are
used, false signals far outnumber the true single transits in any traditional
search.

At the heart of all periodic transit search procedures is a filtering step
based on ``box least squares'' \citep[\bls;][]{Kovacs:2002}.
This step produces a list of candidate transit times that is then vetted to
remove the substantial fraction of false signals using some combination of
automated heuristics and expert curation.
In practice, the fraction of false signals can be substantially reduced by
requiring that at least three self-consistent transits be observed \citep[The
\kepler\ Data Processing Handbook%
\footnote{\url{https://archive.stsci.edu/kepler/manuals/KSCI-19081-001_Data_Processing_Handbook.pdf}};][]{Petigura:2013,
Burke:2014, Rowe:2015, Coughlin:2015}.

Recent work has yielded discoveries of long-period transiting planets with
only one or two transits identified in archival \kepler\ and \KT\ light curves
by visual inspection \citep{Wang:2013, Kipping:2014a, Osborn:2016,
Kipping:2016, Uehara:2016, Wang:2015}.
Since all of these discoveries rely on human interaction, it is intractable to
reliably constrain the completeness of these catalogs and they cannot be used
to measure the occurrence rate of long-period planets.

In this \paper, we develop a systematic method of reliably discovering the
transits of large long-period companions in photometric time series
\emph{without human intervention}.
Since the search methodology is fully-automated, we can robustly measure the
search completeness~--~using injection and recovery tests~--~and use these
products to place probabilistic constraints on the occurrence rate of
long-period planets.
We apply this method to a subset of the archival data from the \kepler\
Mission and produce a catalog of discoveries.

@@

\begin{figure*}
\vspace{\baselineskip}
\begin{center}
\includegraphics[width=\textwidth]{figures/model_comp.pdf}
\end{center}
\caption{%
Representative examples of candidate events flagged by the initial search.
Each example falls into a different model category and the figure shows the
data as black points and the best fit mean model prediction.
The examples represent the following model categories:
\emph{(a)} variability, \emph{(b)} step, \emph{(c)} box, and \emph{(d)}
transit.
\figlabel{model-comp}}
\end{figure*}

@@

\begin{figure}
\vspace{\baselineskip}
\begin{center}
\includegraphics{figures/targets.pdf}
\end{center}
\caption{%
Blah.
\figlabel{targets}}
\end{figure}

@@

\begin{figure}
\vspace{\baselineskip}
\begin{center}
\includegraphics{figures/full_sample.pdf}
\end{center}
\caption{%
Blah.
\figlabel{full-sample}}
\end{figure}

@@

\begin{figure}
\vspace{\baselineskip}
\begin{center}
\includegraphics[width=\textwidth]{figures/completeness.pdf}
\end{center}
\caption{%
\figlabel{completeness}}
\end{figure}

@@

% \acknowledgments
It is a pleasure to thank
\ldots
for helpful contributions to the ideas and code presented here.

DWH was partially supported by the National Science Foundation (grant
IIS-1124794), the National Aeronautics and Space Administration (grant
NNX12AI50G), and the Moore--Sloan Data Science Environment at NYU.

This research made use of the NASA \project{Astrophysics Data System} and the
NASA Exoplanet Archive.
The Exoplanet Archive is operated by the California Institute of Technology,
under contract with NASA under the Exoplanet Exploration Program.
This \paper\ includes data collected by the \kepler\ mission. Funding for the
\kepler\ mission is provided by the NASA Science Mission directorate.
We are grateful to the entire \kepler\ team, past and present.
Their tireless efforts were all essential to the tremendous success of the
mission and the successes of \KT, present and future.

These data were obtained from the Mikulski Archive for Space Telescopes
(MAST).
STScI is operated by the Association of Universities for Research in
Astronomy, Inc., under NASA contract NAS5-26555.
Support for MAST is provided by the NASA Office of Space Science via grant
NNX13AC07G and by other grants and contracts.

Computing resources were provided by High Performance Computing at New York
University.

\facility{Kepler}
\software{%
    \project{ceres} \citep{Agarwal:2016},
    \project{corner.py} \citep{Foreman-Mackey:2016},
    \project{emcee} \citep{Foreman-Mackey:2013},
    \project{george} \citep{Ambikasaran:2016},
	\project{matplotlib} \citep{Hunter:2007},
	\project{numpy} \citep{Van-Der-Walt:2011},
	\project{scipy} \citep{Jones:2001}}.

\newpage
\appendix

@@

\clearpage
\bibliography{peerless}
